Characteristic polynomials of p-adic matrices

### 2017-04-16 charpoly/charpoly.tex
Comments on referees' report

\newcommand{\done}[1]{\textcolor{blue}{#1}}
\newcommand{\tdo}[1]{\textcolor{red}{#1}}

@@

\item \tdo{In section 3, Algorithm 1: what is the cost of step 5, and how did you avoid quartic complexity?}
\item \tdo{In section 4, give a concrete construction for approximating the characteristic polynomial as a ``by product'' of the work of [15].}

@@

\item \tdo{Think about whether Proposition 5.6 can be sped up using the work of Kedlaya and Umans.}

@@

\item \tdo{Convince reviewer that the complexity claimed in Theorem 4.1 is accurate: ``it lacks a precision loss analysis and relies on classic division based algorithms}

@@

\item \tdo{The definitions of companion matrix in the proof of Proposition 2.1 and in Section 2.3 are inconsistent (coefficients on the last column or last row)}

@@

\item \tdo{In the experiments, (section 5.1), it is not clear which of the two methods is being used to compute the comatrix.}

@@

\item \tdo{Make the relationship between $P$, $Q$ and $R$ in Theorem 4.1 explicit.}

@@

\item \tdo{In the abstract, clarify what we mean by ``precision exactly $O(p^k)$.''}
\item \tdo{In the introduction, we say ``Let $M$ be known at precision $O(p^m)$.''  What does this mean?  Coefficient-wise?  Overall?  This needs to be said clearly.}

@@

\item \tdo{Please define lattice in the tangent space and provide a simple, illustrative example.}

@@

@@

\begin{theo}
\label{thm:compute_shortcom}
Given $M \in M_n(K)$ such that $d \chi_M$ is surjective,
then one can compute $(\alpha, P, Q, \chi_M)$
in $K[X]$ such that
$\com(X{-}N) = \alpha \cdot Y^T \cdot Y R^T \mod \chi_M$
in $\softO(n^\omega)$ operations in $K.$
\end{theo}

>>>>

\newcommand{\done}[1]{\textcolor{blue}{#1}}
\newcommand{\tdo}[1]{\textcolor{red}{#1}}
\definecolor{answer}{rgb}{0,0.5,0.2}
\newcommand{\xavier}[1]{\textcolor{answer}{{\bf Xavier:} #1}}
\newcommand{\tristan}[1]{\textcolor{answer}{{\bf Tristan:} #1}}
\newcommand{\david}[1]{\textcolor{answer}{{\bf David:} #1}}

@@

\item \tdo{In section 3, Algorithm 1: what is the cost of step 5, and how did you avoid quartic complexity?}\\
      \xavier{I think that the cost is $O(n^2)$ because it corresponds to an elementary operation of rows/columns. Is that correct?}
\item \tdo{In section 4, give a concrete construction for approximating the characteristic polynomial as a ``by product'' of the work of [15].}\\
      \xavier{I do not really understand this remark. I mean: once we have the Frobenius invariants, the characteristic polynomial is just the product of them, isn't it?}

@@

\item \tdo{Think about whether Proposition 5.6 can be sped up using the work of Kedlaya and Umans.}\\
      \xavier{I cannot see how Kedlaya--Umans can help: the cost $\softO(n^\omega)$ comes from the computation of a Frobenius normal form...}

@@

\item \tdo{Convince reviewer that the complexity claimed in Theorem 4.1 is accurate: ``it lacks a precision loss analysis and relies on classic division based algorithms}\\
      \xavier{One can argue that we do not need \emph{a priori} much accuracy on the precision of $\com(A-X)$ because only the valuations of the entries of the latter matrix are relevant. So counting operations in $K$ makes sense. By the way, we could say at some point that a good idea would be to use relaxed algorithms and increase the precision until we are able to read the valuations.}

@@

      \xavier{Ok pour switching $\com$ to $\text{Adj}$ (and adding a sentence explaining the difference with the previous paper).}
\item \tdo{The definitions of companion matrix in the proof of Proposition 2.1 and in Section 2.3 are inconsistent (coefficients on the last column or last row)}\\
      \xavier{Is that enough to replace ``the bottom row'' by ``the last column'' in the proof of Proposition 2.1?}

@@

      \xavier{FWIW, I wrote a proof in my lecture at JNCF. Maybe we can just cite it.}
\item \tdo{In the experiments, (section 5.1), it is not clear which of the two methods is being used to compute the comatrix.}\\
      \xavier{Ok but it is not important (since we are just interested in the result). Should we really say more about that? (In practice, I think that we used the \texttt{adjoint} method of \textsc{sage}.)}

@@

\item \tdo{Make the relationship between $P$, $Q$ and $R$ in Theorem 4.1 explicit.}\\
      \xavier{\done{Done.} I've rewritten everything in terms of $P$ and $Q$ only.}

@@

\item \tdo{In the abstract, clarify what we mean by ``precision exactly $O(p^k)$.''}\\
      \xavier{By the way, we should also mention in the abstract the $\softO(n^\omega)$ algorithm.}
\item \tdo{In the introduction, we say ``Let $M$ be known at precision $O(p^m)$.''  What does this mean?  Coefficient-wise?  Overall?  This needs to be said clearly.}\\
      \xavier{Well, it is the same, isn't it? Anyway, we may add ``coefficient-wise'' if it's not clear.}

@@

\item \tdo{Please define lattice in the tangent space and provide a simple, illustrative example.}\\
      \xavier{Maybe we should instead remove any mention to the tangent space: it's not really necessary.}

@@

\xavier{I propose to expand the first paragraph of Section 2.1 (and even
maybe make a separate subsection for it). Concretely, I suggest to 
(1)~do not talk about the tangent space (but instead consider lattices
in $M_n(K)$),
(2)~give precision definition of flat, jagged precisions and relate them 
with lattices,
(3)~maybe cite the notes of my recent lecture at JNCF in which all of
this is detailed.}

\bigskip

@@

\begin{theo}
\label{thm:compute_shortcom}
Given $M \in M_n(K)$ such that $d \chi_M$ is surjective,
then one can compute $(\alpha, P, Q, \chi_M)$
in $K[X]$ such that
$\com(X{-}N) = \alpha \cdot P Y^T \cdot Y Q^T \mod \chi_M$
in $\softO(n^\omega)$ operations in $K.$
\end{theo}
